Music Versus No Music Effectiveness On Cognitive Response Time and Typing Efficiency

### 2024-03-30 paper/paper.tex
add related works about music and cognitive function

\author{Shea Spalding}
\email{Shea.Spalding@colostate.edu} %update email 
\affiliation{%
  \institution{Colorado State University}
  \city{Fort Collins}
  \state{Colorado}
  \country{USA}
  \postcode{80525-0650} %chktex 8 %the rest of the zip will need to be corrected
}

@@

% Put the related works here
\section{Related Works}
While there is not necessarily a concrete answer for whether music or background noise is effective on cognitive function, there have been many
studies that attempt to determine the answer. The other studies attempted to search for this answer by
focusing on word processing on a computer with the background music and the user's ability to formulate sentences~\cite{ransdell2001141}, 
or determining if background music help users to keep their attention on their task at hand~\cite{kiss2021effect}.

>>>>

\author{Shea Spalding}
\email{Shea.Spalding@colostate.edu} %update email 
\affiliation{%
  \institution{Colorado State University}
  \city{Fort Collins}
  \state{Colorado}
  \country{USA}
  \postcode{80523-1873} %chktex 8
}

@@

\section{Related Works}
While there is not necessarily a concrete answer for whether music or background noise is effective on cognitive function, there have been many
studies that attempt to determine the answer. The other studies attempted to search for this answer by
focusing on word processing on a computer with the background music and the user's ability to formulate sentences~\cite{ransdell2001141}, 
or determining if background music help users to keep their attention on their task at hand~\cite{kiss2021effect}. Similar to~\cite{kiss2021effect},~\cite{mathew2022inherently} also examined how motivation and music were used hand-in-hand on ``a time-consuming task and task performance''. A different study done in 2010 again found similar results of music both being a hinderance and benefit~\cite{AudioDistractionsAshley}. According to~\cite{Gonzalez_Aiello_2019}, the degree of cognitive function may be based on an individual's personality.

Although~\cite{ransdell2001141} suggests that the music is a detriment, those with ``musical training and high working memory span wrote better essays with longer sentences'. This suggests that while the music slowed down the participants, it did not necessarily lower their cognitive ability in typing on their computer while focusing on forming text for the written essay for this particular study. For~\cite{AudioDistractionsAshley}, the music caused enough of a difference in the time that was taken for writing a paragraph but ``there were no signigicant differences in typing speed between the audio distractions''. This again may suggest that the participants cognitive ability was not hindered when typing and may transfer to directly after should the participants taken a test measuring their cognitive function. The slow timing was again seen in~\cite{Gonzalez_Aiello_2019} in which ``music generally impaired performance on a complex task but improved performance on a simple task''.

The study done through~\cite{kiss2021effect} saw opposite results as to~\cite{Gonzalez_Aiello_2019}. The music or background noise ``enchanced'' the reaction time and performance of the tasks that we given to be completed. The results from~\cite{mathew2022inherently} were split. The music was a detriment for logic puzzles in the user's accuracy but aided in the efficiency of a writing task.
